---
title: "Coursera Practical Machine Learning Project"
author: "Wang"
date: "04/09/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Introduction
### Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

### Data
The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har.

### What you should submit
The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.

### Outcome intepretation
Exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).

Read more: http://groupware.les.inf.puc-rio.br/har#ixzz5yXr4FycK

## Data Processing
```{r message=FALSE, warning=FALSE}
library(caret)
library(randomForest)
library(rpart)
set.seed(1234)
training <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", na.strings = c("#DIV/0!", "NA", ""))
test <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", na.strings = c("#DIV/0!", "NA", ""))
```
### Data Cleaning
```{r}
# Eliminate the irrelevent variables of both training and test set.
training <- training[, -(1:5)]
test <- test[, -c(1:5)]
# Eliminate the variables that include too many missing values (over a threshold of 60% missing values)
sumofNA <- function(x) sum(!is.na(x)) >= 19622*0.4
result <- apply(training, 2, FUN = sumofNA)
training <- training[,names(which(result == TRUE))]
test <- test[,names(which(result == TRUE))[-length(names(which(result == TRUE)))]]
test <- cbind(test, problem_id = c(1:20))
# Eliminate nzv using nearZeroVar function
training <- training[, which(nearZeroVar(training, saveMetrics=TRUE)$nzv == FALSE)]
test <- test[, which(nearZeroVar(test, saveMetrics=TRUE)$nzv == FALSE)]
```

### Spliting the data into training set and testing set. The ratio is set to be 6:4.
```{r}
inTrain = createDataPartition(y=training$classe, p=0.6, list=FALSE)
trainset = training[inTrain,]
testset = training[-inTrain,]
```
## Grow the regression tree model and predict on the test set
```{r}
fit_tree <- rpart(classe~., method = "class", data = trainset)
pred_tree <- predict(fit_tree, testset, type = "class")
confusionMatrix(pred_tree, testset$classe)
```
### K-fold cross_validation
```{r}
crossvalidation <- train(classe~., data = trainset, method = "rpart", trControl = trainControl(method = "cv", number = 10))
crossvalidation
```

## Build the randomForest
```{r}
fit_forest <- randomForest(classe~., data = trainset)
pred_forest <- predict(fit_forest, testset)
confusionMatrix(pred_forest, testset$classe)
# Therefore choose this model (rf)
predict(fit_forest, test, type = "class")
```
## Write the final prediction
```{r}
write.table(predict(fit_forest, test, type = "class"), file = "prediction.txt", quote = FALSE, col.names = c("id"))
```





